Supplementary Content for the article *The role of functional diversity in ecosystem regime shifts*

### 2022-06-02 experiments/2_supplement/supplement.Rmd
requested changes to figures done
changes to supplement

---
title: Supplementary Content for the article *The role of functional diversity in ecosystem regime shifts*
author: "Owen L Petchey, plus all authors of the main article"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
---

@@

plot_ss_result6(repl_method_1,
                temp_meth$ssfind_result[[1]],
                xlims = c(-8, 0))
```

@@

# Previously figure 5 Main Manuscript
```{r fig_5_log, echo = FALSE,  include=TRUE, fig.cap='Previously figure 5 Main Manuscript log', fig.height=8, fig.width=8, fig.align="center"}
## Figure 5 ----
## Assessing additivity of diversity effect sizes
## By Uriah
## Owen switched to SB_tot

@@

fig5_log <- lapply(fig5_log, function(df){
  tag <- tags[as.character(unique(df$var_treat))]
  plot <- df %>%
    filter(Species == "SB_tot", num_strains==9) %>%
    ggplot(aes(x = stand_var, linetype=Method, col=Calculation)) +
    geom_hline(yintercept = c(-8, 0), col = "black", linetype=3) +
    geom_line(aes(y = hyst_min_log), size = 1) +
    geom_line(aes(y = hyst_max_log), size = 1) +
    facet_wrap( ~ var_treat, scales = "free_y", nrow = 3) +
    labs(y="Oxygen diffusivity\n[log10 uM per hour]",
         x="Amount of trait variation\n[see text for units]",
         linetype="Method",col="Calculation", tag=tag) +
    scale_color_manual(values = colours[unique(df$Calculation)]) +
    scale_linetype_manual(values = c(1,2))+
    coord_flip() +
    theme_bw()+
    theme(legend.position="none",
          axis.title = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank())

ylab + Reduce("+",fig5_log) + fig5legend +
  plot_spacer() + xlab + plot_spacer() +
  plot_layout(ncol = 3, heights = c(25,1), widths = c(1,21,3))

@@

```{r fig_5_linear, echo = FALSE,  include=TRUE, fig.cap='Previously figure 5 Main Manuscript linear', fig.height=8, fig.width=8, fig.align="center"}

@@

>>>>

---
title: Supplementary Content for the article *The role of functional diversity in ecosystem regime shifts*
author: "Owen L Petchey, Uriah Daugaard, Anubhav Gupta, Rainer M. Krug, Kimberley D. Lemmen, Sofia J. van Moorsel, Marcel Suleiman, Debra Zuppinger-Dingley, Romana Limberger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
---

@@

plot_ss_result6_new(repl_method_1,
                    temp_meth$ssfind_result[[1]],
                    xlims = c(-8, 0))
```

@@

# (Non-)additivity

We investigated if the effects of trait variation of individual functional groups were additive. We compared the tipping point positions as determined by our simulations with the expected position of the respective tipping point if effects of diversity in different functional groups were additive. The expected tipping point was calculated by adding the effect sizes (i.e. the difference in tipping point position between the no-diversity scenario and a given amount of trait variation) on a linear scale.

```{r fig_5_log, echo = FALSE,  include=TRUE, fig.cap='Previously figure 5 Main Manuscript log', fig.height=8, fig.width=8, fig.align="center"}
## Figure 5 ----
## Assessing additivity of diversity effect sizes
## By Uriah
## Owen switched to SB_tot

@@

# fig5_log <- lapply(fig5_log, function(df){
#   tag <- tags[as.character(unique(df$var_treat))]
#   plot <- df %>%
#     filter(Species == "SB_tot", num_strains==9) %>%
#     ggplot(aes(x = stand_var, linetype=Method, col=Calculation)) +
#     geom_hline(yintercept = c(-8, 0), col = "black", linetype=3) +
#     geom_line(aes(y = hyst_min_log), size = 1) +
#     geom_line(aes(y = hyst_max_log), size = 1) +
#     facet_wrap( ~ var_treat, scales = "free_y", nrow = 3) +
#     labs(y="Oxygen diffusivity\n[log10 uM per hour]",
#          x="Amount of trait variation\n[see text for units]",
#          linetype="Method",col="Calculation", tag=tag) +
#     scale_color_manual(values = colours[unique(df$Calculation)]) +
#     scale_linetype_manual(values = c(1,2))+
#     coord_flip() +
#     theme_bw()+
#     theme(legend.position="none",
#           axis.title = element_blank(),
#           axis.ticks = element_blank(),
#           axis.text = element_blank())
#   
#   if(tag %in% c("a","d")){
#     plot <- plot + theme(axis.text.y = element_text(),
#                          axis.ticks.y = element_line())
#   }
#   if(tag %in% c("h","i")){
#     plot <- plot + theme(axis.text.x = element_text(),
#                          axis.ticks.x = element_line())
#   }
#   if(tag == "g"){
#     plot <- plot + theme(axis.text = element_text(),
#                          axis.ticks = element_line())
#   }
#   plot
# })
# 
# 
# ylab + Reduce("+",fig5_log) + fig5legend +
#   plot_spacer() + xlab + plot_spacer() +
#   plot_layout(ncol = 3, heights = c(25,1), widths = c(1,21,3))

@@

```{r fig_5_linear, echo = FALSE,  include=TRUE, fig.height=8, fig.width=8, fig.align="center"}

@@

(ref:fig_5_linear)  Figure 9. Assessing the additivity or non-additivity of the effects of trait variation. Solid black lines depict the position of the two tipping points at different levels of trait variation, as simulated with the model. The colored, dashed lines show the expected position of tipping points if effects of diversity in different functional groups were additive. The expected tipping point was calculated by adding the effect sizes of trait variation of different functional groups on a linear scale. For example, in (d) the effect size of variation in only CB was added to the effect size of variation in only SB. In (g), the effect size of variation in only CB was added to the effect size of simultaneous variation in both SB and PB.  (CB: cyanobacteria, SB: sulfate-reducing bacteria, PB: phototrophic sulfur bacteria). The dotted vertical lines at -8 and 0 on the x-axis show the extent of the oxygen diffusivity gradient investigated in the simulations.

When two functional groups varied in traits, the effects of trait variation were additive when one of the groups had no diversity effect on a tipping point but non-additive when both groups had diversity effects on a tipping point (Fig. 9 a-f). High trait variation in cyanobacteria and sulfate-reducing bacteria caused the oxic-anoxic tipping point to occur at lower oxygen diffusivity than expected assuming additivity of the individual effects of diversity in these two groups (Fig. 9d). When the two groups of sulfur bacteria varied in traits, the anoxic-oxic tipping point occurred at lower oxygen diffusivity than expected by additivity (Fig. 9f).

When all three groups varied in traits, only the diversity effect in phototrophic sulfur bacteria on the oxic-anoxic tipping point was additive; all other effects were non-additive, especially at higher trait variation (Fig. 9g-i). Both tipping points occurred at lower levels of oxygen diffusivity than expected by additivity. Interestingly, diversity in the cyanobacteria moved the shift to the oxic state to lower levels of oxygen diffusivity although independent variation in the cyanobacteria had no effect on this tipping point.
